Further Modeling of $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak

### 2020-11-05 Papers/Paper2/Submit/Paper.tex
Lastest version of paper2

\begin{abstract}
An improved resonant plasma response model that describes the interaction between a tokamak plasma and a resonant magnetic perturbation  (RMP) is developed. 
The model interpolates between the linear and the nonlinear response regimes, and takes into account the fact that the slip-frequency
is non-zero in the nonlinear regime. The improved model is incorporated into the {\tt EPEC} toroidal asymptotic matching code.
The modified {\tt EPEC} code is used to investigate RMP-induced edge-localized-mode (ELM) mitigation/suppression in DIII-D H-mode discharge \#145380. 
Allowing for a finite slip-frequency (i.e., relaxing the so-called no-slip constraint) is found to slightly facilitate the locking of
driven magnetic island chains to the RMP, and, hence, to slightly facilitate RMP-induced ELM mitigation/suppression.  The previously obtained conclusion 
that the response of a typical H-mode tokamak plasma to an RMP cannot be accurately modeled by linear theory is confirmed. The previously obtained conclusion that the best agreement between theory and observations is achieved by assuming that the
natural frequencies of tearing modes, in the absence of the RMP, are determined by the local equilibrium ${\bf E}\times{\bf B}$ frequency is also confirmed.
\end{abstract}

@@

The recently developed, toroidal, asymptotic matching code {\tt EPEC} (Extended Perturbed Equilibrium Code) has been used to simulate 
RMP-induced ELM mitigation/suppression in DIII-D H-mode plasmas.\cite{rftor,rftor1}
The original {\tt EPEC} model incorporates the no-slip constraint.
The first aim of this paper is to investigate 
what happens when the no-slip constraint is relaxed in the aforementioned {\tt EPEC} simulations.
 The second aim is to examine what happens when an improvement is made to the manner in which the resonant
plasma response component of the {\tt EPEC} model interpolates between the linear and the nonlinear regimes.

@@

\subsection{Plasma Equilibrium}
Consider a large aspect-ratio, low-$\beta$, tokamak plasma whose magnetic flux-surfaces map out (almost)
concentric circles in the poloidal plane.
Such a plasma is well approximated as a periodic cylinder.  Suppose that the minor radius of the plasma
is $a$. Standard cylindrical coordinates ($r$, $\theta$, $z$) are adopted. The system is assumed to
be periodic in the $z$-direction, with periodicity length $2\pi\,R_0$, where $R_0\gg a$ is the simulated
plasma major radius. It is convenient to define the simulated toroidal angle $\phi=z/R_0$.

The equilibrium magnetic field is written ${\bf B} = [0,B_\theta(r), B_\phi]$. The associated equilibrium
plasma current density takes the form ${\bf j} =[0,0,j_\phi(r)]$, where $\mu_0\,j_\phi= (1/r)\,d(r\,B_\theta)/dr$. The
so-called ``safety-factor'', $q(r)=r\,B_\phi/(R_0\,B_\theta)$, parameterizes the helical pitches of equilibrium magnetic
field-lines.

\subsection{Plasma Response to RMP}
Consider the response of the plasma to an externally generated, static, helical, RMP. 
Suppose that the RMP has $m$ periods in the poloidal direction, and $n$ periods in the toroidal direction.
It is convenient to express the perturbed magnetic field and the perturbed plasma current density in
terms of a perturbed poloidal magnetic flux (divided by $2\pi\,R_0$), $\psi(r,\theta,\phi,t)$. In fact, $\delta{\bf B} = \nabla\psi\times \nabla z$, 
and $\mu_0\,\delta {\bf j} = - \nabla^2\psi\,\nabla z$,
where $\psi(r,\theta,\phi,t) = \hat{\psi}(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\phi)]$, and $\hat{\psi}$ is real. This particular representation
is valid provided that $m/n\gg a/R_0$.\cite{rfa}

@@

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that  plasma is trapped inside the
magnetic separatrix of the  island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is usually the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip-frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip-frequency is negligible.)

\subsection{Composite Model}
We can combine Eqs.~(\ref{e3}), (\ref{e4}), (\ref{e5}), and (\ref{e7}) to give the following
composite resonant response model that interpolates between the linear and the nonlinear
regimes:
\begin{align}\label{e9}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} &= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),\\[0.5ex]
\frac{d\varphi_s}{dt}& = \omega_s - \frac{E_{sv}}{\tau_R}\left(\frac{r_s}{({\cal I}/2)\,W_s+\delta_s}\right)\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).\label{e10}
\end{align}
If we define $X_s = |{\mit\Psi}_s|\,\cos\varphi_s$, $Y_s = |{\mit\Psi}_s|\,\sin\varphi_s$, $X_v = |{\mit\Psi}_v|\,\cos\varphi_v$, and $Y_v = |{\mit\Psi}_v|\,\sin\varphi_v$ 
then the previous two equations are more conveniently written in the non-singular (when $|{\mit\Psi}_s|=0$) forms:\,\cite{slip1,slip2,slip3}
\begin{align}\label{e11}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dX_s}{dt}+\omega_s\,Y_s\right)&= E_{ss}\,X_s + E_{sv}\,X_v,\\[0.5ex]
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dY_s}{dt}-\omega_s\,X_s\right)& = E_{ss}\,Y_s + E_{sv}\,Y_v,\label{e12}
\end{align}
where
\begin{equation}\label{e23x}
W_s = 4\left(\frac{\sqrt{X_s^{\,2}+Y_s^{\,2}}}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s.
\end{equation}
With the benefit of hindsight, Eqs.~(\ref{e11}) and (\ref{e12}) could have been derived more directly from the
following heuristic generalization of Eq.~(\ref{e1}):\,\cite{slip2}
\begin{equation}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v.
\end{equation}

@@

Equation~(\ref{e15}) represents a minor improvement to Eq.~(D1) of Ref.~\onlinecite{rftor} in which the interpolation between the
linear and nonlinear regimes is performed in a slightly more accurate manner.

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k\,0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k\,0}= \varpi_{k\,0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k\,0}$
is the so-called ``natural frequency''  (in the absence of the RMP) at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma in response to the RMP.
Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) of Ref.~\onlinecite{rftor} that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.

If we define $X_k = \hat{\mit\Psi}_k\,\cos\varphi_k$ and $Y_k=\hat{\mit\Psi}_k\,\sin\varphi_k$ then Eqs.~(\ref{e15}) and (\ref{e19}) are
more conveniently written in the non-singular (when $\hat{\mit\Psi}_k=0$) forms
\begin{align}\label{e21}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dX_k}{d\hat{t}} 
+\hat{\varpi}_k\,Y_k\right)= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,X_{k'}  -\sin\xi_{kk'}\,Y_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\cos\zeta_k,\\[0.5ex]
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dY_k}{d\hat{t}} -\hat{\varpi}_k\,X_k\right)
= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,Y_{k'}
+\sin\xi_{kk'}\,X_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\sin\zeta_k,\label{e22}
\end{align}
where
\begin{equation}\label{e23}
\hat{W}_k =\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)^{1/2}_{\hat{r}_k} (X_k^{\,2} + Y_k^{\,2})^{\,1/4}.
\end{equation}
Equations (\ref{e15}), (\ref{e19}), (\ref{e21}), (\ref{e22}) and (\ref{e23}) are the toroidal generalizations of Eqs.~(\ref{e9}), (\ref{e10}), (\ref{e11}), (\ref{e12}), and (\ref{e23x}), respectively.

\section{{\tt EPEC} Modeling of DIII-D Discharge \#145380}\label{s4}
\subsection{Introduction}
DIII-D discharge \#145380 is an ITER-Similar-Shape (ISS), ELMing, H-mode discharge, with a toroidal magnetic field $B_T=-1.9\,{\rm T}$, in which the
plasma current, $I_p$, is slowly ramped over a 2 second interval in order to scan the magnetic safety-factor.\cite{d3d,d3d2} The majority ions are Deuterium, whereas the
minority ions are Carbon~VI.
Figure~\ref{fig1} gives an overview of this discharge. ${\mit\Psi}_N$ is the conventional normalized poloidal magnetic flux.
A static RMP is applied to the plasma by running steady $n=3$ currents through the I-coil system.\cite{icoil}
Three windows of ELM suppression or mitigation are evident: the first (in which ELMs are mitigated, but not
entirely suppressed) extends from $t=2840$--$2980$ ms; the
second (in which ELM are entirely suppressed) extends from $t=3320$--$3560$ ms; and the third (in which ELMs are
entirely suppressed) extends from $t=3880$--$4200$ ms. The 
first, second, and third ELM-suppression/mitigation windows extend over the $\overline{q_{95}}$ ranges $3.88$--$3.94$, $3.62$--$3.73$,
and $3.33$--$3.48$, respectively. Here, $\overline{q_{95}}$ is least-squares linear fit to the variation of the safety-factor at the ${\mit\Psi}_N = 0.95$ flux-surface with time.

\subsection{Relaxation of No-Slip Constraint}\label{s4a}
As is discussed in Ref.~\onlinecite{rftor1}, the best agreement between the predictions of the {\tt EPEC} code and experimental observations is
obtained by assuming that the unperturbed natural frequencies at the various rational surfaces in the plasma are determined by the
local equilibrium ${\bf E}\times {\bf B}$ frequencies. In other words, 
\begin{equation}\label{e24}
\varpi_{k\,0} =-n\,\omega_E({\mit\Psi}_{p\,k}),
\end{equation}
where 
\begin{equation}
\omega_E({\mit\Psi}_p) = -\frac{d{\mit\Phi}}{d{\mit\Psi}_p}.
\end{equation}
Here, ${\mit\Phi}({\mit\Psi}_p)$ is the equilibrium scalar electric potential (in the absence of the RMP), ${\mit\Psi}_p$ the equilibrium poloidal magnetic flux (divided by $2\pi$), 
and ${\mit\Psi}_{p\,k}$  the value of ${\mit\Psi}_p$ at the $k$th rational surface.  Equation~(\ref{e24}) holds at
every rational surface in all of the simulations discussed in Sect.~\ref{s4a}.

@@

The upper panel of Fig.~\ref{fig2} shows the natural frequencies, in the absence of the RMP, of the various  $n=3$ tearing modes 
 that are resonant in the pedestal of DIII-D discharge \#145380. It can be seen that 
the natural frequencies of the $m=11$ (brown), the $m=10$ (magenta), and the $m=9$ (cyan), tearing modes have their
zero-crossings in the  first, second, and third, ELM-suppression/mitigation window, respectively. 
  The lower panel of Fig.~\ref{fig2} shows the natural frequencies, in the presence of the RMP, of the various $n=3$
tearing modes that are resonant in the pedestal. It is apparent that if the magnitude of the natural frequency in the absence of the RMP falls below about
$3\,{\rm krad/s}$ then the associated tearing mode locks to the RMP (i.e., its true natural frequency becomes zero).

The upper panel of Fig.~\ref{fig3} shows the widths of the various vacuum magnetic island chains driven by the applied $n=3$ RMP in the pedestal region as functions of $
\overline{q_{95}}$.
The vacuum island widths are the  expected island widths in the absence of any shielding of driven magnetic reconnection due to
plasma flow; in other words, they are the  widths of the island chains that would be driven if all of the natural frequencies were zero. It can be seen that
the driven magnetic island chains extend over most of the pedestal, implying a very significant degradation in pedestal energy and particle confinement due to the
expected flattening of the density and temperature profiles across the chains.\cite{flat} Moreover, the driven island chains at the
bottom of the pedestal overlap, implying that the magnetic field in this region is rendered stochastic. The lower panel
of Fig.~\ref{fig3} shows the predicted widths of the various magnetic island chains driven by the applied $n=3$ RMP as functions of time. It is apparent 
that the predicted island widths are generally much smaller than the vacuum island widths, as a consequence of the shielding of driven magnetic
reconnection due to plasma flow. However, the shielding breaks down at various rational surfaces when the associated natural frequency is zero (see the lower panel of
Fig.~\ref{fig3}). In this situation, the driven island width is similar to the vacuum island width.

According to the prevailing hypothesis, RMP-induced ELM mitigation/suppression in an H-mode tokamak discharge occurs when a comparatively wide, locked, magnetic
island chain is generated at the top of the pedestal.\cite{hu} The temperature and density flattening associated with the chain
is presumed to limit the radial expansion of the
pedestal, and, thereby, prevent it from attaining a width sufficient to destabilize peeling-ballooning modes\,\cite{connor}  (which are thought to trigger ELMs).\cite{d3d}
It has also been suggested that  the well-known ``density pump-out" phenomenon, which is often observed when RMPs are applied
to H-mode tokamak plasmas, is associated with the formation of overlapping island chains
at the bottom of the pedestal.\cite{hu}

According to Fig.~\ref{fig3},  
 the comparatively wide, $m=11$ (brown),  $m=10$ (magenta), and  $m=9$ (cyan), locked magnetic island chains which are predicted to be driven by the RMP in the pedestal of DIII-D discharge \#145380 line up
fairly well with the first, second, and third, ELM-suppression/mitigation windows, respectively. To be more exact, the island chains are located at
the top of the pedestal (i.e., ${\mit\Psi}_N=0.925$), and also occur in roughly the correct time intervals. This is consistent with the hypothesis that the formation of locked
magnetic island chains at the top of the pedestal causes ELM mitigation/suppression. 
Note that overlapping island chains
are also driven at the bottom of the pedestal; as has already been mentioned, such island chains are associated with the density pump-out phenomenon.

@@

A comparison of Figs.~\ref{fig2} and \ref{fig4}, as well as an examination of Fig.~\ref{fig6}, reveals that the relaxation of the
no-slip constraint facilitates the locking of magnetic island chains to the RMP to a relatively modest degree. (Here, by locking, we mean a reduction in the
RMP-modified natural frequency to a value close to zero.) 
This is not surprising, since
relaxation of the no-slip constraint clearly weakens the coupling between a given magnetic island chain and the local plasma flow (which is ultimately responsible
for the shielding of driven magnetic reconnection). 
It can be seen from Fig.~\ref{fig6} that if the no-slip constraint is imposed then the RMP-modified natural frequencies of locked magnetic
island chains are exactly zero. On the other hand, if the no-slip constraint is relaxed then locked magnetic island chains possess
 small but finite RMP-modified natural frequencies, because the local plasma flow is able to slip through the  stationary chains to some
 extent. However, the degree of slippage for a fully developed locked magnetic island chain is comparatively small (i.e., no more than
 $0.5\,{\rm krad/s}$), which helps explain
 why the no-slip constraint can be used to accurately predict the dynamics of wide magnetic island chains in tokamaks and reversed field pinches.\cite{rfp}

>>>>

\begin{abstract}
An improved resonant plasma response model that describes the interaction between a tokamak plasma and a resonant magnetic perturbation  (RMP) is developed. 
The model interpolates between the linear and the nonlinear response regimes, and takes into account the fact that the slip-frequency
is non-zero in the nonlinear regime. The improved model is incorporated into the {\tt EPEC} toroidal asymptotic matching code.
The modified {\tt EPEC} code is used to investigate RMP-induced edge-localized-mode (ELM) suppression in DIII-D H-mode discharge \#145380. 
Somewhat surprisingly, allowing for a finite slip-frequency (i.e., relaxing the so-called no-slip constraint) is found to only slightly facilitate the locking of
driven magnetic island chains to the RMP, and, hence, to only slightly facilitate RMP-induced ELM suppression.  The previously obtained conclusion 
that the response of a typical H-mode tokamak plasma to an RMP cannot be accurately modeled by linear theory is confirmed. The previously obtained conclusion that the best agreement between theory and observations is achieved by assuming that the
natural frequencies of tearing modes, in the absence of the RMP, are determined by the local equilibrium ${\bf E}\times{\bf B}$ frequency is also confirmed.
\end{abstract}

@@

The recently developed, toroidal, asymptotic matching code {\tt EPEC} (Extended Perturbed Equilibrium Code) has been used to simulate 
RMP-induced ELM suppression in DIII-D H-mode plasmas.\cite{rftor,rftor1}
The original {\tt EPEC} model incorporates the no-slip constraint.
The first aim of this paper is to investigate 
what happens when the no-slip constraint is relaxed in the aforementioned {\tt EPEC} simulations.
 The second aim is to examine what happens when an improvement is made to the manner in which the resonant
plasma response component of the {\tt EPEC} model interpolates between the linear and the nonlinear regimes.

@@

\subsection{Plasma Equilibrium}
Consider a large aspect-ratio, low-$\beta$, circular cross-section, tokamak plasma.
Such a plasma is well approximated as a periodic cylinder.  Suppose that the minor radius of the plasma
is $a$. Standard cylindrical coordinates ($r$, $\theta$, $z$) are adopted. The system is assumed to
be periodic in the $z$-direction, with periodicity length $2\pi\,R_0$, where $R_0\gg a$ is the simulated
plasma major radius. It is convenient to define the simulated toroidal angle $\phi=z/R_0$. 
The equilibrium magnetic field is written ${\bf B} = [0,B_\theta(r), B_\phi]$.  The
so-called ``safety-factor'', $q(r)=r\,B_\phi/(R_0\,B_\theta)$, parameterizes the helical pitches of equilibrium magnetic
field-lines.

\subsection{Plasma Response to RMP}
Consider the response of the plasma to an externally generated, static, helical, RMP. 
Suppose that the RMP has $m$ periods in the poloidal direction, and $n$ periods in the toroidal direction.
It is convenient to express the perturbed magnetic field and the perturbed plasma current density in
terms of a perturbed poloidal magnetic flux (divided by $2\pi\,R_0$), $\psi(r,\theta,\phi,t)$. In fact, $\delta{\bf B} = \nabla\psi\times \nabla z$, 
where $\psi(r,\theta,\phi,t) = \hat{\psi}(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\phi)]$, and $\hat{\psi}$ is real. This particular representation
is valid provided that $m/n\gg a/R_0$.\cite{rfa}

@@

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that  plasma is trapped inside the
magnetic separatrix of the  island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is usually the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.

\subsection{Composite Model}
If we define $X_s = |{\mit\Psi}_s|\,\cos\varphi_s$, $Y_s = |{\mit\Psi}_s|\,\sin\varphi_s$, $X_v = |{\mit\Psi}_v|\,\cos\varphi_v$, and $Y_v = |{\mit\Psi}_v|\,\sin\varphi_v$ 
then we can combine Eqs.~(\ref{e3}), (\ref{e4}), (\ref{e5}), and (\ref{e7}) to give the following
composite resonant response model that interpolates between the linear and the nonlinear
regimes:\,\cite{slip1,slip2,slip3}
\begin{align}\label{e11}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dX_s}{dt}+\omega_s\,Y_s\right)&= E_{ss}\,X_s + E_{sv}\,X_v,\\[0.5ex]
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dY_s}{dt}-\omega_s\,X_s\right)& = E_{ss}\,Y_s + E_{sv}\,Y_v,\label{e12}
\end{align}
where
\begin{equation}\label{e23x}
W_s = 4\left(\frac{\sqrt{X_s^{\,2}+Y_s^{\,2}}}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s.
\end{equation}
With the benefit of hindsight, Eqs.~(\ref{e11}) and (\ref{e12}) could have been derived more directly from the
following heuristic generalization of Eq.~(\ref{e1}):\,\cite{slip2}
\begin{equation}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v.
\end{equation}

@@

Equation~(\ref{e15}) represents a {\em minor}\/ improvement to Eq.~(D1) of Ref.~\onlinecite{rftor} in which the interpolation between the
linear and nonlinear regimes is performed in a slightly more accurate manner.

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k\,0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k\,0}= \varpi_{k\,0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k\,0}$
is the so-called ``natural frequency''  (in the absence of the RMP) at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma in response to the RMP.

Equation~(\ref{e19}) represents
a {\em major}\/ improvement to Eq.~(D6) of Ref.~\onlinecite{rftor} that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.

If we define $X_k = \hat{\mit\Psi}_k\,\cos\varphi_k$ and $Y_k=\hat{\mit\Psi}_k\,\sin\varphi_k$ then Eqs.~(\ref{e15}) and (\ref{e19}) are
more conveniently written in the non-singular (when $\hat{\mit\Psi}_k=0$) forms
\begin{align}\label{e21}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dX_k}{d\hat{t}} 
+\hat{\varpi}_k\,Y_k\right)= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,X_{k'}  -\sin\xi_{kk'}\,Y_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\cos\zeta_k,\\[0.5ex]
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dY_k}{d\hat{t}} -\hat{\varpi}_k\,X_k\right)
= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,Y_{k'}
+\sin\xi_{kk'}\,X_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\sin\zeta_k,\label{e22}
\end{align}
where
\begin{equation}\label{e23}
\hat{W}_k =\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)^{1/2}_{\hat{r}_k} (X_k^{\,2} + Y_k^{\,2})^{\,1/4}.
\end{equation}
Equationsv (\ref{e21}), (\ref{e22}) and (\ref{e23}) are the toroidal generalizations of Eqs.~(\ref{e11}), (\ref{e12}), and (\ref{e23x}), respectively.

\section{{\tt EPEC} Modeling of DIII-D Discharge \#145380}\label{s4}
\subsection{Introduction}
DIII-D discharge \#145380 is an ITER-Similar-Shape (ISS), ELMing, H-mode discharge in which the
plasma current is slowly ramped in order to scan the magnetic safety-factor.\cite{d3d,d3d2} This discharge is described in more detail in Sect.~III.B of Ref.~\onlinecite{rftor1}.
A static RMP is applied to the plasma by running steady $n=3$ currents through the I-coil system.\cite{icoil}
Three windows of ELM suppression  are evident: the first occurs when $t\simeq 2900$ ms and $q_{95}\simeq 3.91$; the
second occurs when $t=3400$ ms and $q_{95}\simeq 3.66$; and the third occurs when $t\simeq 4000$ ms and $q_{95}\simeq 3.41$.  Here, $q_{95}$ is the safety-factor at the 95\% flux-surface.

\subsection{Relaxation of No-Slip Constraint}\label{s4a}
As is discussed in Ref.~\onlinecite{rftor1}, the best agreement between the predictions of the {\tt EPEC} code and experimental observations is
obtained by assuming that the unperturbed natural frequencies at the various rational surfaces in the plasma are determined by the
local equilibrium ${\bf E}\times {\bf B}$ frequencies. In other words, 
\begin{equation}\label{e24}
\varpi_{k\,0} =-n\,\omega_E({\mit\Psi}_{p\,k}),
\end{equation}
where 
$\omega_E({\mit\Psi}_p) = -d{\mit\Phi}/d{\mit\Psi}_p$.
Here, ${\mit\Phi}({\mit\Psi}_p)$ is the equilibrium scalar electric potential (in the absence of the RMP), ${\mit\Psi}_p$ the equilibrium poloidal magnetic flux (divided by $2\pi$), 
and ${\mit\Psi}_{p\,k}$  the value of ${\mit\Psi}_p$ at the $k$th rational surface.  Equation~(\ref{e24}) holds at
every rational surface in all of the simulations discussed in Sect.~\ref{s4a}.

@@

The upper panel of Fig.~\ref{fig2} shows the natural frequencies, in the absence of the RMP, of the various  $n=3$ tearing modes 
 that are resonant in the pedestal of DIII-D discharge \#145380, whereas  
  the lower panel of Fig.~\ref{fig2} shows the same frequencies  in the presence of the RMP.
   It is apparent that if the magnitude of the natural frequency in the absence of the RMP falls below about
$3\,{\rm krad/s}$ then the associated tearing mode locks to the RMP (i.e., its true natural frequency becomes zero).

The upper panel of Fig.~\ref{fig3} shows the widths of the various  magnetic island chains driven by the applied $n=3$ RMP in the
absence of shielding due to plasma flow, whereas the lower panel
shows the same widths in the presence of shielding. It is apparent 
that the predicted island widths are generally much smaller than the ``vacuum'' island widths (i.e., the widths in the absence of shielding), as a consequence of the shielding of driven magnetic
reconnection by plasma flow. However, this shielding breaks down at various rational surfaces when the associated natural frequency is zero (see the lower panel of
Fig.~\ref{fig3}). In this situation, the driven island width is similar to the vacuum island width.

According to the prevailing hypothesis, RMP-induced ELM mitigation/suppression in an H-mode tokamak discharge occurs when a comparatively wide, locked, magnetic
island chain is generated at the top of the pedestal.\cite{hu} The temperature and density flattening associated with the chain
is presumed to limit the radial expansion of the
pedestal, and, thereby, prevent it from attaining a width sufficient to destabilize peeling-ballooning modes\,\cite{connor}  (which are thought to trigger ELMs).\cite{d3d}

According to Fig.~\ref{fig3},  
 the comparatively wide, $m=11$ (brown),  $m=10$ (magenta), and  $m=9$ (cyan), locked magnetic island chains which are predicted to be driven by the RMP in the pedestal of DIII-D discharge \#145380 line up
fairly well with the first, second, and third, ELM-suppression windows, respectively. To be more exact, the island chains are located at
the top of the pedestal (i.e., ${\mit\Psi}_N=0.925$), and also occur in roughly the correct time intervals. This is consistent with the hypothesis that the formation of locked
magnetic island chains at the top of the pedestal causes ELM mitigation/suppression.

@@

A comparison of Figs.~\ref{fig2} and \ref{fig4}, as well as an examination of Fig.~\ref{fig6}, reveals that the relaxation of the
no-slip constraint facilitates the locking of magnetic island chains to the RMP to a relatively modest degree. (Here, by locking, we mean a reduction in the
RMP-modified natural frequency to a value close to zero.) 
This is not surprising, because 
relaxation of the no-slip constraint clearly weakens the coupling between a given magnetic island chain and the local plasma flow (which is ultimately responsible
for the shielding of driven magnetic reconnection). 
It can be seen from Fig.~\ref{fig6} that if the no-slip constraint is imposed then the RMP-modified natural frequencies of locked magnetic
island chains are exactly zero. On the other hand, if the no-slip constraint is relaxed then locked magnetic island chains possess
 small but finite RMP-modified natural frequencies, because the local plasma flow is able to slip through the  stationary chains to some
 extent. However, the degree of slippage for a fully developed locked magnetic island chain is comparatively small (i.e., no more than
 $0.5\,{\rm krad/s}$), which helps explain
 why the no-slip constraint can be used to accurately predict the dynamics of wide magnetic island chains in tokamaks and reversed field pinches.\cite{rfp}
